---
title: "Visualisations of which agents use specific groups of action"
subtitle: "Killing"
author:
- Jill Walker Rettberg, University of Bergen, jill.walker.rettberg@uib.no
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    number_sections: yes
    code_folding: 
    df_print: paged
  pdf_document:
    toc: yes

editor_options:
  markdown:
    wrap: 72
---

This notebook contains scripts that allow you to set a group of actions
(verbs in the dataset), to find synonyms of those actions that are also
in the database, and to create plots to show what kinds of agents use
those verbs.

# Setup

First we'll load the packages we're using.

```{r, setup, message=FALSE, echo=FALSE}
library(tidyverse)
library(syn)
```

## Set the actions 

In the next bit of code, we define the actions we're interested in.

```{r define_target_actions, message=TRUE, echo=TRUE}

target_action <- c("Killing", "Killed")

# Define what you want the actions to be called in the plot.

name_of_target_actions <- "Killing-related actions"
name_of_NON_target_actions <- "Not killing-related"

plot_title <- "Which agents are killing or killed?"

```

Now import situations.csv.

```{r import_situations}
Situations <- read_csv("../data/situations.csv",
                       col_types = cols(
                               SituationID = col_integer(),
                               Situation = col_character(), 
                               Genre = col_character(),
                               Character = col_character(),
                               Entity = col_character(),
                               Technology = col_character(),
                               Verb = col_character()
                       )
)
```

## Look for synonyms

Let's see if there are any more actions we should be including as well
by using the syn() function to find synonyms to the verbs we set in
target_action. Note that the syn package is based on the Moby thesaurus
which is from 1996 and is often weak on verbs - it has none of the
"Hacking" verbs in it, for instance, but many "Flawed" verbs.

```{r find_synonyms}
syns(target_action)

```

When `syn()` returns `character(0)` that means that there are no
synonyms for the word in the thesaurus. If there are synonyms, you can
run the following code to redefine target_actions to include both the
verbs you entered at the start and the synonyms. The next lines convert
the verbs to title case and delete any that don't end in -ing or -ed and
finally those that aren't in our dataset.

```{r add_synonyms}

target_action <- c(target_action, unlist(lapply(target_action, syn)))
target_action <- str_to_title(
        target_action[str_detect(target_action, "ing")|str_detect(target_action, "ed")])
target_action <- target_action[target_action %in% Situations$Verb]

# Show the actions it'll use - maybe these aren't the best and need tweaking?
target_action

# If these words aren't good, just manually define target_action to include only the 
# actions you want, using the following syntax:
# target_action <- c("Verb1", "Verb2")
```

# Visualisations

Now we're ready to start examining the results.

## Which agents perform the target actions?

```{r plot_of_agents_and_target_actions, fig.height = 7}

Situations %>% 
        mutate(Target_action = case_when(Verb %in% target_action ~ name_of_target_actions,
                                  TRUE ~ name_of_NON_target_actions),
               Agent = case_when(!is.na(Entity) ~ "Entity",
                                 !is.na(Technology) ~ "Technology",
                                 !is.na(Character) ~ "Character")) %>% 
        filter(Target_action == name_of_target_actions) %>% 
        select(Genre, Verb, Entity, Technology, Agent) %>% 
        pivot_longer(!Verb, values_to = "value") %>%
        drop_na() %>% 
        ggplot(aes(x=fct_infreq(value), fill=Verb)) +
        geom_bar(aes(y = ..count..))+
        theme(axis.line = element_line(colour = "darkblue", 
                                       size = 1, 
                                       linetype = "solid")) +
        theme_minimal() +
        labs(title = plot_title,
             subtitle="",
             x = "") +
        coord_flip() +
        facet_wrap(~name, scales="free") 
```

## Same visualisation for agents that are NOT related to killing or being killed

```{r plot_of_agents_NOT_doing_target_actions, fig.height = 7}

Situations %>% 
        mutate(Target_action = case_when(Verb %in% target_action ~ name_of_target_actions,
                                         TRUE ~ name_of_NON_target_actions),
               Agent = case_when(!is.na(Entity) ~ "Entity",
                                 !is.na(Technology) ~ "Technology",
                                 !is.na(Character) ~ "Character"),
               VerbTense = case_when(str_detect(Verb, "ing") ~ "Active",
                                     TRUE ~ "Passive")) %>% 
        filter(Target_action != name_of_target_actions) %>% 
        select(Genre, Verb, Entity, Technology, Agent, VerbTense) %>% 
        pivot_longer(!c(Verb, VerbTense), values_to = "value") %>%
        drop_na() %>% 
        ggplot(aes(x=fct_infreq(value), fill = VerbTense)) +
        geom_bar(aes(y = ..count..))+
        theme(axis.line = element_line(colour = "darkblue", 
                                       size = 1, 
                                       linetype = "solid")) +
        theme_minimal() +
        labs(title = "Distribution of agents in all the OTHER actions",
             subtitle="",
             x="") +
        coord_flip() +
                facet_wrap(~name, scales="free") 
```

## Compare the target actions to the rest of the actions in the dataset

The next plot compares the target actions - in this case actions
relating to hkilling and being killed by machine vision - to the rest of the
dataset. This is a "facet_grid" plot that shows a separate pane for each
genre by whether the verbs are active or passive. Each of these six
panes shows which agents, proportionally, do hacking-related and
non-hacking related actions.

```{r}

Situations %>% 
        mutate(Target_action = case_when(Verb %in% target_action ~ name_of_target_actions,
                                         TRUE ~ name_of_NON_target_actions),
               Agent = case_when(!is.na(Entity) ~ "Entity",
                                 !is.na(Technology) ~ "Technology",
                                 !is.na(Character) ~ "Character"),
               VerbTense = case_when(str_detect(Verb, "ing") ~ "Active",
                                     TRUE ~ "Passive")) %>% 
       #filter(Verb != "Drained") %>%  # remove # to manually remove any verbs that seem wrong
        select(Target_action, Agent, Genre, VerbTense) %>% 
        group_by(Agent, Target_action) %>% 
        ggplot(aes(x = Target_action, fill = Agent)) +
        geom_bar(position="fill") +
        labs(title = "Actions shown by genre and whether they are active (-ing) or passive (-ed)",
             subtitle="",
             x = "Actions",
             y = "Proportion of all interactions in dataset") +
        coord_flip() +
        facet_grid(VerbTense~Genre)

```

This is quite a dense visualisation to interpret!

## Same plot but by count instead of proportion

The last plot was proportional, so you don't notice that of course the
vast majority of actions is not related to the target actions. This is
the same plot, but showing absolute numbers instead of the proportions.

```{r by_count_instead_of_prop, fig.width=7}

Situations %>% 
        mutate(Target_action = case_when(Verb %in% target_action ~ name_of_target_actions,
                                         TRUE ~ name_of_NON_target_actions),
               Agent = case_when(!is.na(Entity) ~ "Entity",
                                 !is.na(Technology) ~ "Technology",
                                 !is.na(Character) ~ "Character"),
               VerbTense = case_when(str_detect(Verb, "ing") ~ "Active",
                                     TRUE ~ "Passive")) %>% 
       #filter(Verb != "Drained") %>%  # remove # to manually remove any verbs that seem wrong
        select(Target_action, Agent, Genre, VerbTense) %>% 
        group_by(Agent, Target_action) %>% 
        ggplot(aes(x = Target_action, fill = Agent)) +
        geom_bar() +
        labs(title = "Actions shown by genre and whether they are active (-ing) or passive (-ed)",
             subtitle="",
             x = "Actions",
             y = "Proportion of all interactions in dataset") +
        coord_flip() +
        facet_grid(VerbTense~Genre)

```
